% Remerciements / Acknowledgements
%
%  Grâce aux remerciements, l'auteur attire l'attention du lecteur
% sur l'aide que certaines personnes lui ont apportée, sur leurs
% conseils ou sur toute autre forme de contribution lors de la
% réalisation de son mémoire. Le cas échéant, c'est dans cette section
% que le candidat doit témoigner sa reconnaissance à son directeur de
% recherche, aux organismes dispensateurs de subventions ou aux
% entreprises qui lui ont accordé des bourses ou des fonds de
% recherche.
\ifthenelse{\equal{\Langue}{english}}{
	\chapter*{ACKNOWLEDGEMENTS}\thispagestyle{headings}
	\addcontentsline{toc}{compteur}{ACKNOWLEDGEMENTS}
}{
	\chapter*{REMERCIEMENTS}\thispagestyle{headings}
	\addcontentsline{toc}{compteur}{REMERCIEMENTS}
}
I was introduced to Swarm Intelligence research in 2019 when Professors Giovanni Beltrame, David St-Onge and Nicolas Reeves gave me the opportunity to work on the creation of an artistic and technological installation for Chambord Castle's $500^{th}$ anniversary. This sparked my interest for distributed systems and research, and I would like to thank them for their trust which later led me to pursuing a master's degree.

This work and my master's would not have been possible without the great help and guidance from Professor Giovanni Beltrame. Throughout the last two year, he taught me invaluable research and engineering skills. I greatly appreciate his patience and availability to answer all my questions.  

I was lucky enough to work with David Vielfaure on many of my projects, and I thank him for his insights, collaboration and support.

I would also like to thank Jérôme Collin and Benjamin De Leneer for trusting me respectively as their laboratory teacher assistant and their lecturer. Through my passion for teaching, they allowed me to transmit my interest of computer engineering to future generations of engineers.

I must also to thank Polytechnique Montréal, \href{https://www.groupedeschenes.com}{Groupe Deschênes} and the \href{https://www.quebec.ca/en/government/ministere/enseignement-superieur}{Ministry of Higher Education} for providing financial support which allowed me to pursue my master's degree. 

Finally, I want to thank my friends and family for their support and encouragements throughout the last years.
